Weakly informative observations: not the winner

### 2020-08-23 inst/notthewinner.Rmd
slight rewriting
works with names paradigm

Here we consider a situation in which we make an observation that is
only very very weakly informative.  We consider the observation
${\mathcal O}$ that a particular player *does not* win: the order
statistic is unobserved and we know nothing about it except for the
fact that (wlog) player 1 ($p_1$, aka "me") is not first placed.  We
consider whether there is evidence to support two different assertions:

@@

Our alternative would be that $(p_1,\ldots,p_n) is unconstrained
except for the unit sum constraint:

@@

```{r}
(H <- hyper2(pnames=paste("p",1:5,sep="")) +
    rank_likelihood(c(5,2,1,4)) + 
    rank_likelihood(c(2,4,1)) + 
    rank_likelihood(c(4,3,1)) + 
    rank_likelihood(c(3,5)))
	```

@@

```{r}
n <- 5 # number of players
small <- 1e-3  # numerical necessity
(mH1 <- maxp(H,startp=c(small*2,rep(small,n-2)),fcm=rbind(c(1,0,0,-1)),fcv=0))
loglik(indep(mH1),H)
loglik(indep(mH0),H)-loglik(indep(mH1),H)   # support
```

@@

```{r,uninformativesupportfunction}
library("partitions")
H # as before, here for convenience
M <- perms(5)   # uninformative!
SL <- list()
for(i in seq_len(ncol(M))){
	SL[[i]] <- H + rank_likelihood(M[,i,drop=TRUE])
}
SL <- as.suplist(SL)
```

@@

```{r,mls0,cache=TRUE}
mSL0 <- maxp(SL)
mSL1 <- maxp(SL,startp=c(small*2,rep(small,n-2)),fcm=rbind(c(1,0,0,-1)),fcv=0)
mSL2 <- equalp(SL)
```

@@

# Additional weakly informative observation

Now we make another observation ${\mathcal O}$ which is independent of
${\mathcal I}$: person `1` did not win.  It is clear that observation
${\mathcal O}$ is only very weakly informative.  Package idiom is
similar to the perfectly uninformative observation considered above,
except that we remove every order that has person 1 winning:

```{r,slacalc}
M <- perms(5)
M <- M[, M[1,]!=1]  # remove any column with "1" in first place
SLa <- list()
for(i in seq_len(ncol(M))){
	SLa[[i]] <- H + rank_likelihood(M[,i,drop=TRUE])
}
SLa <- as.suplist(SLa)
```

So `SLa` is a compound object that gives a support function for the
joint observation ${\mathcal I}$ and ${\mathcal O}$.  We can find the
MLE:

```{r,mlsa,cache=TRUE}
mSLa0 <- maxp(SLa)
mSLa1 <- maxp(SLa,startp=c(small*2,rep(small,n-2)),fcm=rbind(c(1,0,0,-1)),fcv=0)
mSLa2 <- equalp(SLa)
```

```{r,suppacalc}
Suppa <-
      log(c(
      H0 = like_single_list(indep(mSLa0),SLa),
      H1 = like_single_list(indep(mSLa1),SLa),
      H2 = like_single_list(indep(mSLa2),SLa)))
(Suppa <- Suppa-max(Suppa))
```

We can now compare the support function for ${\mathcal I}$ with the
support function for ${\mathcal I}$ and ${\mathcal H}$:

```{r,comparesups}
Supp2-Suppa
```

Thus ${\mathcal O}$ furnishes a tiny bit of extra evidence against
$H_1$ (about $0.016$ units of support) and a little bit of extra
evidence against $H_2$ (about $0.193$ units of support).

>>>>

(see also `inst/loser.Rmd`).
Here we consider a situation in which we make
an observation that is only very very weakly informative.  We consider
the observation ${\mathcal O}$ that a particular player *does not*
win: the order statistic is unobserved and we know nothing about it
except for the fact that (wlog) player 1 ($p_1$, aka "me") is not
first placed.  We consider whether there is evidence to support two
different assertions:

@@

Our alternative would be that $(p_1,\ldots,p_n)$ is unconstrained
except for the unit sum constraint:

@@

```{r}
(H <- 
	rank_likelihood(c(a=3,c=2,b=1,d=4)) + 
	rank_likelihood(c(b=2,c=3,d=1    )) + 
	rank_likelihood(c(d=2,a=3,c=1    )) + 
	rank_likelihood(c(b=1,d=2        )) )
```

@@

```{r}
samep.test(H,c(1,4))
```

@@

```{r,uninformativesupportfunction}
library("partitions")
H # as before, here for convenience
M <- perms(5)   # uninformative!
SL <- list()
for(i in seq_len(ncol(M))){
	jj <- M[,i,drop=TRUE]
	names(jj) <- letters[seq_along(jj)]
	SL[[i]] <- H + rank_likelihood(jj)
}
SL <- as.suplist(SL)
```

@@

```{r,mls0,cache=TRUE}
n <- 5 # number of players
small <- 1e-4
mSL0 <- maxplist(SL)
mSL1 <- maxplist(SL,startp=c(small*2,rep(small,n-2)),fcm=rbind(c(1,0,0,-1)),fcv=0)
mSL2 <- equalp(SL)
```

@@
